---
title: "Data Wrangling"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## This document provides helpful tips for data wrangling using R

#### Importing the required libraries

The libraries can be imported by using the function library. The suppressmessage() function suppresses the message displayed while a library is loaded.

```{r }
suppressMessages(library(readr)) 
suppressMessages(library(dplyr))
suppressMessages(library(tidyr))
```


#### Basic data explorations commands


```{r}
x <- c(20,10,32,17,73,15, 22, 35, 41, 58, 36, 7, 8, 49, 20, 11, 1, 2, 123, 135, 34, 5)
head(x) # returns the first few lines of the data
tail(x) # returns the last few lines of the data
nrow(x) # return the number of rows
ncol(x) # return the number of column
```


#### Generic data generation commands


```{r}
df1 = read_csv("../Data/sample_file2.txt", col_names = TRUE)
c(1:10)# Concatenate function to create a vector with numbers from 1 to 10.
seq(0, 100, 10) # Produces a sequence from 0 to 100 in steps of 10.
rep(2,10)# replicates 2, 10 times. 
rep(c(1,2,3),2) #outputs 1 2 3 1 2 3. 
rep(c("a","b","c"), each=2) # gives a a b b c c.
selCols = df1[,c("ID","AMOUNT")] #sub-sets the variables ID and YIELD
df1[1:7, 1:3] #returns the first 7 rows and the first 3 columns of data frame df1.
df2=df1[20:25,] #returns a subset of rows 20 to 25.
cleanData=df1[-c(2,7,15),] #removes rows 2, 7 and 15.
selRows=df1[df1$VAR1==3,] #sub-sets with rows where VAR1 = 3.
df2=transform(df1, VAR2=VAR1*2) #. Creates variable VAR2 that is twice the value of VAR1
```


#### tidyr and dplyr - The packages commonly used for data wrangling in R. 

Some of these notes below have been created using https://www.rstudio.com/wp-content/uploads/2015/02/data-wrangling-cheatsheet.pdf

```{r iris}
glimpse(iris) # displays the first few rows of data
View(iris) # displays first few rows of data in a spreadsheet like tabular format. V is capitalized
summary(iris) # Provides the summary of the data including mean, median, min and max values
count(iris,Species) # Counts for each value in the column specified
#gather is used to convert column into rows based on the values and therefore elongating the table while reducing the width of the table
#spread is opposite of gather. It converts the columns to rows therefore reduces the length of the table while increasing the width
filter(iris, Petal.Length==1.3) #filters the data based on the condition provided
distinct(iris) # removes duplicates
sample_n(iris, 5, replace = FALSE) # select 5 rows at random
slice(iris, 3:8) # extract 3rd to 8th rows
group_by(iris,Petal.Length) # returns a table grouped by the col specified.
iris   %>%   group_by(Species)   %>%   rename(Type = Species) %>% glimpse() #piping function %>% Passes the values from the left(could be a df, variable or another function) to the function on the right for further processing
```