---
title: "Final Report - An analysis of Bike Vendors data"
author: "Vishal Bhalla"
date: "April 22, 2019"

output: 
  html_document:
    fig_caption: true
    toc: true
    
---
&nbsp;

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r message=FALSE, warning=FALSE, echo=FALSE}
# First of all import the libraries that will be required for the complete excercise
library(dplyr) # provides piping operator
library(janitor) # provides functions to clean the data and headers
library(tidyverse) #data manipulation functions
library(cluster) #clustering algorithms
library(reshape2) #for melt function
```
&nbsp;

####  1. Introduction
This document contains the analysis done on the **Bike Vendors** data set. The report is being prepared as part of the final submission for the *Applied Data Science* class.

First we load the data into a Data Frame and then analyse its features. We then formulate our questions that we need to get the answers for. Thereafter we use the techniques we learnt to get those answers from the data provided.

```{r message=FALSE, warning=FALSE}
bike_raw <- read_csv("../Data/bikeVendors.csv") # Read the data
```
&nbsp;

####  2. Questions that need to be answered
 * Which bike sells more? Mountain Bike or Road Bike?
 * Which store sells most bikes?
 * What is the cost of a bike?


&nbsp;

####  3. Understand the data

```{r, eval=TRUE, include=FALSE}
head(bike_raw)
glimpse(bike_raw) # Display the number of observations and variables in the data frame. Display few records of the data set.
bike <- bike_raw %>%
        clean_names() # Removing extra column spaces
sapply(bike, function(x) sum(is.na(x)))
bike %>%
  select(-price) %>%
    summarize_if(is.numeric, sum)
bike1 = bike # make a copy of the df
```
&nbsp;

We see that there are **35** variables and **97** obeservations for each variable. There are no missing values.

There are 4 character variables namely **model, category1, category2, and price**. The remaining 31 variables are of type double, of which one is **price** and rest of the variables appear to be the names of the vendors. 

The variable names are displayed below.
```{r echo=FALSE}
as.data.frame(colnames(bike))
```


&nbsp;

The **model** variable contains the name of the bike model, the **category1** variable contain the category of the bike, i.e. if it is a Mountain Bike or a Road bike and **category2** further sub categorize it by usage like Sports or Elite Road bike for example.  The **Frame** variable contains the details on the type of frame 'Carbon' or 'Aluminium' and **Price** contains the price of the frame.


The data under each vendor name has a sum total of 1, so it appears to be the percentage of each type of frame sold by the vendor totalling to 100%.

&nbsp;

Looking at the data, I think more appropriate questions to ask would be:

 * What is the percentage market share for bikes based on frame type?
 * What is the percentage market share for Mountain Bike and Road Bike?
 * What is the price range where most of the bikes sell?

&nbsp;

####  4. Analysis - Sales
```{r fig.height=5, fig.width=9, echo=FALSE}
bike1 %>%
  select(-model, - category1, -category2,-price) %>%
  group_by(frame) %>%
  summarise_if(is.numeric, sum) %>%
  melt(id.var='frame') %>%
  ggplot(aes(x=variable, y=value, fill=frame)) +
    geom_col() +
    labs(title = "Frame Types sales per Vendor" , y=("% of Sales") , x=("Vendors")) +
    scale_fill_brewer(palette = "Set2") + 
    theme(plot.title=element_text(size=12, hjust=.5), axis.text.x = element_text(face="italic", size=10, angle=90)) +
    geom_hline(yintercept=.5, color="yellow", size=1.5)
```

&nbsp;

From the above figure we can see that the sales of both the carbon as well as aluminium frame is almost equal across the vendors. The yellow line marks 50% of the sales.

&nbsp;


```{r fig.height=5, fig.width=9, echo=FALSE}
bike$totalbyrow <- rowSums(bike[,6:35], na.rm=TRUE)
totalbycat <- bike %>%
  select(category1, category2,totalbyrow) %>%
  group_by(category1, category2) %>%
  summarise_if(is.numeric, sum)
  ggplot(totalbycat, aes(x=category2, y=(totalbyrow*100/35), color=category1)) + geom_point(shape=19,size=3)  +
    labs(title = "Sales share per sub category" , y=("Percentage of Total Sales") , x=("Sub Categories"), color="Types of bike") +
    scale_fill_brewer(palette = "Dark2") + 
    theme(plot.title=element_text(size=12, hjust=.5), axis.text.x = element_text(face="italic", size=10, angle=90))
```

&nbsp;

From the chart above we see that the highest selling bike is the Elite Road bike and it contributes to around 25% of total bike sales.

&nbsp;

```{r fig.height=5, fig.width=9, echo=FALSE}
ggplot(bike, aes(sample = (totalbyrow*100),colour = factor(category1))) + 
  geom_qq(distribution = qnorm) + 
  stat_qq_line() + ylab("Sales %") + labs(color='Types of Bikes', title = "Sales distribution across the vendors") + 
  theme(plot.title=element_text(size=12, hjust=.5))
```

&nbsp;

The sales of the bikes across the various vendors seems to follow almost a normal distribution with some outliers at both the lower and higher edges of the data.

&nbsp;

####  5. Analysis - Price

```{r fig.height=5, fig.width=9, echo=FALSE}
ggplot(bike, aes(x = category1, y = price)) +
  geom_boxplot(fill="firebrick", color="black") + 
  scale_x_discrete(name = "Bike Categories") +
  scale_y_continuous(name = "Price") +
  ggtitle("Price of bike by categories per frame type") + 
  theme(plot.title=element_text(size=12, hjust=.5)) +
  facet_grid(. ~ frame)
```

&nbsp;

The box plots above show that the Carbon frame bikes are expensive when compared to the Aluminum frame bikes. The bikes that fall under the Mountain bike category are expensive than the road bikes. The most expensive bikes are the Mountain Bikes with Carbon frame with the median price aroud $6000.

&nbsp;

```{r fig.height=5, fig.width=9,echo=FALSE}
ggplot(data=bike, aes(price)) + 
  geom_histogram(fill="firebrick",binwidth = 1000, col=("black")) +
  ggtitle("Price Distribution") + 
  theme(plot.title=element_text(size=12, hjust=.5)) +
  geom_vline(aes(xintercept=mean(price), color = "Mean"),linetype="dashed", size=1, show.legend = T) +
  geom_vline(aes(xintercept=median(price), color = "Median"),linetype="dashed", size=1, show.legend =  T) + 
  scale_color_manual("Bike Price", values = c("Mean" = "royalblue", "Median" = "green")) +
  scale_x_continuous(breaks = seq(0, max(bike$price), by = 1000))
```

&nbsp;

The histogram above shows the price distribution for the bikes. We observe that the major number of bikes sold are less than the overall mean price of the bikes. Majority of the bikes range between the price of $500 - $4500.

&nbsp;

```{r fig.height=5, fig.width=9, echo=FALSE}
bike2 <- bike %>%
  select(-model, - category1, -frame,-price) %>%
  group_by(category2) %>%
  summarise_if(is.numeric, sum)
head(bike2)
```

