---
title: "Midterm"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Midterm Exam

This document has been created to present the code which we write as part of our midterm examination for Applied Data Science course

## Import neccessary libraries

```{r }
library(stringr)
library(htmlwidgets)
library(readr)
library(dplyr)
library(tidyverse)
library(nycflights13)
library(ggplot2)
```


## Question 1 - To find the number of columns that have NaN

```{r }
epaData <- read_csv("../data/US EPA data 2017.csv")
```

```{r }
head(epaData)
tail(epaData)
sort(colSums(is.na(epaData)), decreasing = T)
```

## Question 2 - To find the maximum negative correlation between 2 variables

```{r mtcars}
mtc <- dplyr::select(mtcars, -c(cyl, gear, carb, am, vs))
head(mtc)
corrplot::corrplot(cor(mtc),method = "number")
```

## Question 4 - Mean, Trimmed Mean, Weighted Mean, Median, Weighted Median

```{r mtcars}
mean(mtcars$mpg)
mean(mtcars$mpg,trim=.1)
weighted.mean(mtcars$mpg,mtcars$wt)
median(mtcars$mpg)
median(rep(mtcars$mpg, times=mtcars$wt))
```

## Question 7 - SD, IQR and MAD(Median Absolute Deviation)

```{r mtcars}
sd(mtcars$mpg)
IQR(mtcars$mpg)
mad(mtcars$mpg)
```
## Question 8 - Regex match 
### /(\"\w+\":\"\w*\")/im (\"\w+\":\"\d{4}-\d{2}-\d{2}\")

## Question 12
```{r who}
nrow(who)
ncol(who)
head(who)
summary(who)
```
## Question 13
```{r mpg}
ggplot(data = mpg,mapping=aes(displ,hwy,))+
  geom_point(color='Black',fill='Red',shape=24, size=3)
```

## Question 16
```{r}
test <-matrix(1:9, ncol=3)
print(test[2, c(1,3)])
```
## Question 17
```{r faithful}
quantile(faithful$eruptions, probs = .25)
```

## Question 18
```{r iris}
iris%>%
  filter(Petal.Length >4.5)%>%
  group_by(Species)%>%
  summarise(mean(Sepal.Length),sd(Sepal.Width),max(Petal.Length))
```

## Question 19
```{r diamonds}
diamonds%>%
  group_by(color)%>%
  filter(color == 'D')%>%
  summarise(count = n())
```

## Question 20
### using nycflights13 
```{r flights}
flights%>%
  group_by(month)%>%
  filter(month == 1 | month == 6)%>%
  summarise(sum = n())
```

