---
title: "ANOVA"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Working on k sample problems where the sample sizes are same.

###*H0 = mu1 = mu2 = muk*
###*H1 = H0 is not true*

Assumptions

1. Assume Independence

2. Assume/Check Normal Population using QQPLOT

3. Assume/Check equal variances or Homoscedacity


We set a seed value so that the random numbers generated are the same

```{r}
set.seed(12345)
```
Now we generate the random numbers and see if we are getting the numbers with same mean and variance. We also test with a different mean and see how the box plots look in both the cases
```{r}
par(mfrow=c(1,2))
X1 = rnorm(50,mean=50, sd=15)
X2 = rnorm(50,mean=50, sd=15)
X3 = rnorm(50,mean=50, sd=15)
boxplot(X1,X2,X3)
Y1 = rnorm(50,mean=50, sd=15)
Y2 = rnorm(50,mean=50, sd=15)
Y3 = rnorm(50,mean=60, sd=15)
boxplot(Y1,Y2,Y3)
```

We check the variances and we take the average of the 3 variances also called the "Within sample". 

```{r}
var(X1)
var(X2)
var(X3)
within_ms=(var(X1)+var(X2)+var(X3))/3
within_ms
```

Next we look at "Between samples estimate"" of variance which is nothing but the variance of the 3 means multiplied by the number of samples .

```{r}
between_ms=var(c(mean(X1),mean(X2),mean(X3)))*50
var(c(mean(Y1),mean(Y2),mean(Y3)))*50
between_ms
```

Next we calculate F statistics(Fischer Statistics) which is the ratio of within and between values. Then we use the F value to calculate the p value which is then used to either accpet or reject the NULL hypothesis. df1 is the number of samples(3 in our case) and df2 is n value for each sample (50 in this case)
```{r}
F = between_ms/within_ms
F
p = 1 - pf(F,df1 = 2, df2 = 147)
p
```

Next we see how all the above calulcations can be acheived using one command - *Anova* in R
```{r}
x.all = c(X1,X2,X3)
group = factor(c(rep(1,50),rep(2,50),rep(3,50)))
head(x.all)
head(group)
anova(lm(x.all ~ group))
```
