---
title: "Week6"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Import the required libraries


```{r }
library(tidyverse)
```

## Explore mpg

```{r }
str(mpg)
```
## Plot mpg
```{r }
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = class, shape = drv ) )
```
## Plot mpg using Stroke
```{r }
ggplot(data = mpg, aes(x = displ, y = hwy )) + 
  geom_point(shape = 21, size = 3, fill = "blue", stroke = 3, colour = "black" )

```
## Facet Wrap and Facet Grid - 
### Facet using continuous var created large number of grids while a plot with 2 discrete variables places values only at junctions
```{r }
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = drv)) + 
  facet_wrap(~ class, ncol = 2)

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = drv), size = 2) + 
  facet_grid(fl ~ cyl)
```
## Trying out facets
```{r }
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) +
  geom_smooth(mapping = aes(x = displ, y = hwy, linetype = drv, color = drv) )
```
## Trying out the global mappings
```{r }
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point(mapping = aes(color = class)) + 
  geom_smooth(mapping = aes(linetype = drv))
```

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point() + 
  geom_smooth(se = FALSE)

ggplot(data = mpg, mapping = aes(x = displ, y = hwy, group = drv)) + 
  geom_point() + 
  geom_smooth(se = FALSE)

ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = drv)) + 
  geom_point() + 
  geom_smooth(se = FALSE)

ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point(mapping = aes(color = drv)) + 
  geom_smooth(se = FALSE)

ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point(mapping = aes(color = drv)) + 
  geom_smooth(mapping = aes(linetype = drv),se = FALSE)

ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point(mapping = aes(color = drv))
```
## Explore the diamonds dataset
```{r}
str(diamonds)
```
## Testing various geom options
```{r}
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, fill = cut))
ggplot(data = diamonds) + 
  geom_col(mapping = aes(x = cut, y = y, color = cut))
ggplot(data = diamonds) + 
  geom_smooth(mapping = aes(x = carat, y = price, color = cut))
```
## Testing the porps/groups
```{r}
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, y = ..prop.., group = 1))
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, fill = clarity))
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, fill = clarity), position = "fill")
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, fill = clarity), position = "dodge")
```
## Box Plots and coord_flip
```{r}
ggplot(data = diamonds, mapping = aes(x = cut, y = price)) + 
  geom_boxplot()
ggplot(data = diamonds, mapping = aes(x = cut, y = price)) + 
  geom_boxplot() +
  coord_flip()
ggplot(data = mpg, mapping = aes(x = class, y = hwy )) + 
  geom_boxplot()
ggplot(data = mpg, mapping = aes(x = class, y = hwy)) + 
  geom_boxplot() +
  coord_flip()
```
## Working with Polar Co-ordinates
```{r}
bar <- ggplot(data = diamonds) + 
  geom_bar(
    mapping = aes(x = cut, fill = cut), 
    show.legend = FALSE,
    width = 1
  ) + 
  theme(aspect.ratio = 1) +
  labs(x = NULL, y = NULL)

bar + coord_flip()
bar + coord_polar()

bar1 <- ggplot(diamonds,aes(x = cut)) + 
  geom_bar(width = 1, colour = "black"
  )  +
  theme(aspect.ratio = 1) +
  labs(x = NULL, y = NULL)
bar1 + coord_polar()
bar2 <- ggplot(diamonds,aes(x = cut), fill = cut) + 
  geom_bar(width = 1
  )  +
  theme(aspect.ratio = 1) +
  labs(x = NULL, y = NULL)
bar2
+ coord_polar(theta = "y")
```

## Maps
```{r}

nz <- map_data("nz")

ggplot(nz, aes(long, lat, group = group)) +
  geom_polygon(fill = "white", colour = "black")

ggplot(nz, aes(long, lat, group = group)) +
  geom_polygon(fill = "white", colour = "black") +
  coord_quickmap()

ggplot(nz, aes(long, lat, group = group)) +
  geom_polygon(fill = "white", colour = "black") +
  coord_map()

```

```{r}
ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) +
  geom_point() + 
   geom_abline() +
  coord_fixed()
```