---
title: "Final Report - An analysis of Bike Vendors data"
author: "Vishal Bhalla"
date: "April 22, 2019"

output: 
  html_document:
    fig_caption: true
    toc: true
    
---
&nbsp;

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r message=FALSE, warning=FALSE, echo=FALSE}
# First of all import the libraries that will be required for the complete excercise
library(dplyr) # provides piping operator
library(janitor) # provides functions to clean the data and headers
library(tidyverse) #data manipulation functions
library(cluster) #clustering algorithms
library(reshape2) #for melt function
library(factoextra) #clustering algorithms & visualization
library(cluster) #clustering algorithms
```
&nbsp;

####  1. Introduction
This document contains the analysis done on the **Bike Vendors** data set. The report is being prepared as part of the final submission for the *Applied Data Science* class.

First we load the data into a Data Frame and then analyse its features. We then formulate our questions that we need to get the answers for. Thereafter we use the techniques we learnt to get those answers from the data provided.

```{r message=FALSE, warning=FALSE}
bike_raw <- read_csv("../CodePortfolio/R/Data/bikeVendors.csv") # Read the data
```
&nbsp;

####  2. Questions that need to be answered
 * Which bike sells more? Mountain Bike or Road Bike?
 * Which store sells most bikes?
 * What is the cost of a bike?


&nbsp;

####  3. Understand the data

```{r, eval=TRUE, include=FALSE}
head(bike_raw)
glimpse(bike_raw) # Display the number of observations and variables in the data frame. Display few records of the data set.
bike <- bike_raw %>%
        clean_names() # Removing extra column spaces
sapply(bike, function(x) sum(is.na(x)))
bike %>%
  select(-price) %>%
    summarize_if(is.numeric, sum)
bike1 = bike # make a copy of the df
```
&nbsp;

We see that there are **35** variables and **97** obeservations for each variable. There are no missing values.

There are 4 character variables namely **model, category1, category2, and price**. The remaining 31 variables are of type double, of which one is **price** and rest of the variables appear to be the names of the vendors. 

The variable names are displayed below.
```{r echo=FALSE}
as.data.frame(colnames(bike))
```


&nbsp;

The **model** variable contains the name of the bike model, the **category1** variable contain the category of the bike, i.e. if it is a Mountain Bike or a Road bike and **category2** further sub categorize it by usage like Sports or Elite Road bike for example.  The **Frame** variable contains the details on the type of frame 'Carbon' or 'Aluminium' and **Price** contains the price of the frame.


The data under each vendor name has a sum total of 1, so it appears to be the percentage of each type of frame sold by the vendor totalling to 100%.

&nbsp;

Looking at the data, I think more appropriate questions to ask would be:

 * What is the percentage market share for bikes based on frame type?
 * What is the percentage market share for Mountain Bike and Road Bike?
 * What is the price range where most of the bikes sell?

&nbsp;

####  4. Analysis - Sales
```{r fig.height=5, fig.width=9, echo=FALSE}
bike1 %>%
  select(-model, - category1, -category2,-price) %>%
  group_by(frame) %>%
  summarise_if(is.numeric, sum) %>%
  melt(id.var='frame') %>%
  ggplot(aes(x=variable, y=value, fill=frame)) +
    geom_col() +
    labs(title = "Frame Types sales per Vendor" , y=("% of Sales") , x=("Vendors")) +
    scale_fill_brewer(palette = "Set2") + 
    theme(plot.title=element_text(size=12, hjust=.5), axis.text.x = element_text(face="italic", size=10, angle=90)) +
    geom_hline(yintercept=.5, color="yellow", size=1.5)
```

&nbsp;

From the above figure we can see that the sales of both the carbon as well as aluminium frame is almost equal across the vendors. The yellow line marks 50% of the sales.

&nbsp;


```{r fig.height=5, fig.width=9, echo=FALSE}
bike$totalbyrow <- rowSums(bike[,6:35], na.rm=TRUE)
totalbycat <- bike %>%
  select(category1, category2,totalbyrow) %>%
  group_by(category1, category2) %>%
  summarise_if(is.numeric, sum)
  ggplot(totalbycat, aes(x=category2, y=(totalbyrow*100/35), color=category1)) + geom_point(shape=19,size=3)  +
    labs(title = "Sales share per sub category" , y=("Percentage of Total Sales") , x=("Sub Categories"), color="Types of bike") +
    scale_fill_brewer(palette = "Dark2") + 
    theme(plot.title=element_text(size=12, hjust=.5), axis.text.x = element_text(face="italic", size=10, angle=90))
```

&nbsp;

From the chart above we see that the highest selling bike is the Elite Road bike and it contributes to around 25% of total bike sales.

&nbsp;

```{r fig.height=5, fig.width=9, echo=FALSE}
ggplot(bike, aes(sample = (totalbyrow*100),colour = factor(category1))) + 
  geom_qq(distribution = qnorm) + 
  stat_qq_line() + ylab("Sales %") + labs(color='Types of Bikes', title = "Sales distribution across the vendors") + 
  theme(plot.title=element_text(size=12, hjust=.5))
```

&nbsp;

The sales of the bikes across the various vendors seems to follow almost a normal distribution with some outliers at both the lower and higher edges of the data.

&nbsp;

####  5. Analysis - Price

```{r fig.height=5, fig.width=9, echo=FALSE}
ggplot(bike, aes(x = category1, y = price)) +
  geom_boxplot(fill="firebrick", color="black") + 
  scale_x_discrete(name = "Bike Categories") +
  scale_y_continuous(name = "Price") +
  ggtitle("Price of bike by categories per frame type") + 
  theme(plot.title=element_text(size=12, hjust=.5)) +
  facet_grid(. ~ frame)
```

&nbsp;

The box plots above show that the Carbon frame bikes are expensive when compared to the Aluminum frame bikes. The bikes that fall under the Mountain bike category are expensive than the road bikes. The most expensive bikes are the Mountain Bikes with Carbon frame with the median price aroud $6000.

&nbsp;

```{r fig.height=5, fig.width=9,echo=FALSE}
ggplot(data=bike, aes(price)) + 
  geom_histogram(fill="firebrick",binwidth = 1000, col=("black")) +
  ggtitle("Price Distribution") + 
  theme(plot.title=element_text(size=12, hjust=.5)) +
  geom_vline(aes(xintercept=mean(price), color = "Mean"),linetype="dashed", size=1, show.legend = T) +
  geom_vline(aes(xintercept=median(price), color = "Median"),linetype="dashed", size=1, show.legend =  T) + 
  scale_color_manual("Bike Price", values = c("Mean" = "royalblue", "Median" = "green")) +
  scale_x_continuous(breaks = seq(0, max(bike$price), by = 1000))
```

&nbsp;

The histogram above shows the price distribution for the bikes. We observe that the major number of bikes sold are less than the overall mean price of the bikes. Majority of the bikes range between the price of $500 - $4500.

&nbsp;

####  6. Analysis - Cluster

```{r echo=FALSE}
bike2 <- subset(bike, select=c("price", "cincinnati_speed", "detroit_cycles","wichita_speed"))
bike2scaled <-scale(bike2[,-1])
fviz_nbclust(bike2scaled, kmeans, method = "gap_stat")
km4 <- kmeans(bike2scaled, 4, nstart = 30)
fviz_cluster(km4, data = bike2[,-1],
             ellipse.type = "norm",
             palette = "Dark",
             show.clust.cent = TRUE,
             geom = c("point"),
             main = " Cluster Analysis for Vendors cincinnati_speed,detroit_cycles,wichita_speed",
             shape = 19,
             outlier.color = black, 
             outlier.shape = 1,
             ggtheme = theme_classic())
bikesil <- silhouette(km4$cluster, dist(bike2scaled))
fviz_silhouette(bikesil,
                title = " Silhouette Plot for Vendors cincinnati_speed,detroit_cycles,wichita_speed",
                pallete="Dark",
                print.summary = FALSE)
message("Cluster Size:" )
km4$size
```

&nbsp;

As the number of vendors is large we selected 3 vendors randomly("cincinnati_speed", "detroit_cycles","wichita_speed") and performed the kmeans clustering on the same. 

The 4 clusters formed are clear but overlapping. This indicates that the data in itself is not contain natural clusters. The same can be interpreted with the Silhoutte plot where we see that all the clusters are having around 50% affect i.e. the observations fall under 2 clusters. There are even negative observations suggesting that the points are categorized under incorrect clusters. 

&nbsp;

####  7. Analysis - Regression


```{r echo=FALSE}
bike$totalbyrow <- rowSums(bike[,6:35], na.rm=TRUE)
message("Mean of Sales Total:" )
mean(bike$totalbyrow)
message("Mean of Price:" )
mean(bike$price)
par(mfrow=c(2, 2))  # divide graph area in 2 columns
boxplot(bike$price, main="Price")  # box plot for 'price'
boxplot(bike$totalbyrow, main="Sales Totals")  # box plot for 'Sales Totals'
plot(density(bike$price), main="Density Plot: Price", ylab="Frequency")  # density plot for 'price'
polygon(density(bike$price), col="red")
plot(density(bike$totalbyrow), main="Density Plot: Sales Totals", ylab="Frequency")  # density plot for 'Sales Totals'
polygon(density(bike$totalbyrow), col="red")
```
&nbsp;

We perform the regression analysis of the bike vendors data, with the sum of each row as our dependant variable and the price of the bike as independent variable. 

The boxplot and density plots above show that the "Sales Total" for each row of the data is normally distributed and is shaped almost like a bell curve. The density plot for "Price" shows that it is skewed towards the right. The boxplot for price shows that there are few outliers.

The formula we use here is **lm(totalbyrow~price, data=bike)**

The NULL Hypothesis here is that the sales of the bikes have a strong relationship with price of the bike. If we get a small p value it would mean that our NULL Hypothesis should be rejected. 

&nbsp;
```{r echo=FALSE}
lnmdl <- lm(totalbyrow~price, data=bike)
summary(lnmdl)
```
&nbsp;

The residuals seem to be almost symmetrical around a mean value of 0. This wwould suggest that the model is a good fit. It predicts the values very close to the real observations.

The low p co-efficient means that a relationship does not exist between the two variable ie the price and the sum of sales% across all vendors.

A larger F statistics means the NULL hypothesis is true while the low p value points to the NULL Hypothesis not being true. Here the p value is extremely small, so our NULL Hypothesis should be rejected, i.e. there does not exist a relation between the sum sales and the price. 

&nbsp;


```{r echo=FALSE}
par(mfrow = c(2,2))
plot(lnmdl)
```
&nbsp;

The residuals vs fitted graph does not show any particuar pattern. This would mean that there is no non-linear relationship between the predicted and actual observations that would have resulted in the data points being left out.

The residuals almost follow a normal distribution which is a good sign for the linear model. There are some outliers at both the ends though.

The scale location line is almost horizontal, which means that the residuals are well distributed across the range of observations which is a good sign.

There are no outliers beyong the cooks like(dashed red line) which would mean that there are no outliers which have affect on the regression if they are omitted. Therefore we can ignore those observations.

All the four plots show observations that point to a **good linear model fit**.

&nbsp;

####  8. Conclusion

&nbsp;

From the various plots that we generated and observations of the data we can say that the sales of the bikes across the vendors is fairly consistent in terms of the type of frame and also in term of the bike category. We see that almost 50% of the bikes that are sold are that of Aluminum while the other 50% are carbon frame bikes(refer to plot "Frame Types sales per Vendor"). Similarly the plot "Sales share per sub category" shows that the sales across the two categories i.e. road bike and mountain bike is equally distributed amongst the various vendors.

The boxplots "Price of bike by categories per frame type" provide us the visualalization that the Carbon frame bikes are expensive when compared to the Aluminum frame bikes. The "Price Distibution" histogram presents the facts that majority of the bike modesl are available in the price range of $500 - $4500. The mean price of a bike is ~$4000 while the median price is $3250. This difference in mean and median means that there are outliers in the data, which would be the expensive Carbon framed bikes.

&nbsp;

